### 2024-10-02 paper/related_literature.tex
more on related work

@@

\subsection{Data Augmentation for Generalization}

\citet{sauer2021counterfactual} generate counterfactual images for MNIST and ImageNet through independent mechanisms (IM): each IM learns class-conditional input distributions over a specific lower-dimensional, semantically meaningful factor, such as \textit{texture}, \textit{shape} and \textit{background}. The demonstrate that using these generated counterfactuals during classifier training improves model robustness. Similarly, \citet{abbasnejad2020counterfactual} argue that counterfactuals represent potentially useful training data in machine learning, especially in supervised settings where inputs may be reasonably mapped to multiple outputs. They, too, demonstrate the augmenting the training data of image classifiers can improve generalization. \citet{teney2020learning} propose an approach using counterfactuals in training that does not rely on data augmentation: they argue that counterfactual pairs typically already exist in training datasets. Specifically, their approach relies on, firstly, identifying similar input samples with different annotations and, secondly, ensuring that the gradient of the classifier aligns with the vector between pairs of counterfactual inputs using the cosine distance as a loss function (referred to as \textit{gradient supervision}) (\textcolor{red}{this might be useful for our task as well}).

>>>>

\subsection{Background on Counterfactual Explanations}

\subsection{Learning Representations}

\textcolor{red}{For example, joint-energy models ...}

\subsection{Generalization and Robustness}

\citet{sauer2021counterfactual} generate counterfactual images for MNIST and ImageNet through independent mechanisms (IM): each IM learns class-conditional input distributions over a specific lower-dimensional, semantically meaningful factor, such as \textit{texture}, \textit{shape} and \textit{background}. The demonstrate that using these generated counterfactuals during classifier training improves model robustness. Similarly, \citet{abbasnejad2020counterfactual} argue that counterfactuals represent potentially useful training data in machine learning, especially in supervised settings where inputs may be reasonably mapped to multiple outputs. They, too, demonstrate the augmenting the training data of image classifiers can improve generalization. \citet{teney2020learning} propose an approach using counterfactuals in training that does not rely on data augmentation: they argue that counterfactual pairs typically already exist in training datasets. Specifically, their approach relies on, firstly, identifying similar input samples with different annotations and, secondly, ensuring that the gradient of the classifier aligns with the vector between pairs of counterfactual inputs using the cosine distance as a loss function (referred to as \textit{gradient supervision}) (\textcolor{red}{this might be useful for our task as well}). In the natural language processing (NLP) domain, counterfactuals have similarly been used to improve models through data augmentation: \citet{wu2021polyjuice}, propose POLYJUICE, a general-purpose counterfactual generator for language models. They demonstrate empirically that augmenting training data through POLYJUICE counterfactuals improves robustness in a number of NLP tasks.

\subsection{Link to Adversarial Training}

\citet{freiesleben2022intriguing} propose two definitional differences between Adversarial Examples (AE) and Counterfactual Explanations (CE): firstly, and more importantly according to the authors, the term AE implies missclassification, which is not the case for CE (\textcolor{red}{this might be a useful notion for use to distinguish between adversarials and explanations during training}); secondly, they argue that closeness plays a more critical role in the context of CE but confess that even counterfactuals that are not close might be relevant explanations. \citet{pawelczyk2022exploring} show that CE and AE are equivalent under certain conditions and derive upper bounds on the distances between them.

\subsection{Closely Related}

@@

\citet{luu2023counterfactual} introduce Counterfactual Adversarial Training (CAT) with intention of improving generalization and robustness of language models. Specifically, they propose to proceed as follows: firstly, identify training samples that are subject to high predictive uncertainty (entropy); secondly, generate counterfactual explanations for those samples; and, finally, finetune the model on the augmented dataset that includes the generated counterfactuals.
